---
title: "introduction-to-modperformr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction-to-modperformr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ModPerformR)
```

## Introduction
ModPerformR is a package designed for modeling and analyzing running performance. It provides a set of functions to calculate various aspects of running performance, including power output, critical power models, and utilities for working with running data.

## Péronnet-Thibault Model
The Péronnet-Thibault model is used to estimate power output for a given running duration.

```{r}
# Example usage
power_output <- peronnet_thibault_model(
  t = 600,  # Running duration in seconds
  A = 1500,     # Anaerobic capacity in J/kg
  MAP = 25,     # Maximal aerobic power in W/kg
  E = -1.5,     # Endurance in W/kg/log(time in s)
  Tmap = 420,   # Running duration eliciting MAP in seconds
  BMR = 1.2     # Basal metabolic rate in W/kg
)

print(paste("Estimated power output:", round(power_output, 2), "W/kg"))

```

## Critical Power Model

The Critical Power model is another approach to estimate sustainable power output.

```{r}
# Example usage
power_output_cp <- critical_power_model(
  t = 300,     # Duration in seconds
  cp = 300,    # Critical power in watts
  Wprime = 20000 # Anaerobic work capacity in joules
)

print(paste("Estimated power output (Critical Power model):", round(power_output_cp, 2), "W"))

```

## Kennelly Models

The Kennelly models estimate running velocity and power based on distance.

```{r}
distance <- 1000
ck <- 10

velocity <- Kennelly_1906_V(distance, ck)
power <- Kennelly_1906_P(distance, ck)

print(paste("Estimated velocity from Kennelly model:", round(velocity, 2), "m/s"))
print(paste("Estimated power from Kennelly model:", round(power, 2), "W"))

```

## Riegel Models
The Riegel models estimate running time, velocity, and power based on distance.

```{r}
distance <- 1000
a <- 10
b <- 1.06

time <- Riegel_time(distance, a, b)
velocity <- Riegel_velocity(distance, a, b)
power <- Riegel_power(distance, a, b)

print(paste("Time estimated from Riegel model:", round(time, 2), "s"))
print(paste("Velocity estimated from Riegel model:", round(velocity, 2), "m/s"))
print(paste("Power estimated from Riegel model:", round(power, 2), "W"))

```

## Rumball Models
The Rumball models estimate running velocity and power based on distance.

```{r}
distance <- 1000
f <- 0.5
c <- 2

velocity <- Rumball_velocity(distance, f, c)
power <- Rumball_power(distance, f, c)

print(paste("Velocity estimated from Rumball model:", round(velocity, 2), "m/s"))
print(paste("Power estimated from Rumball model:", round(power, 2), "W"))

```

## Lloyd Model
The Lloyd model estimates power output based on time and various physiological parameters.

```{r}
pmax <- 25
g <- 0.05
map <- 25
k <- 10
time <- 100

power <- Lloyd(pmax, g, map, k, time)

print(paste("Power estimated from Lloyd model:", round(power, 2), "W/kg"))

```


## Ward-Smith Model
The Ward-Smith model estimates power output based on time and various physiological parameters.

```{r}
pmax <- 25
map <- 25
lambda <- 0.05
time <- 100

power <- WardSmith(pmax, map, lambda, time)

print(paste("Power estimated from Ward-Smith model:", round(power, 2), "W/kg"))

```

## Omni-Domain Power-Duration Model
The OmniPD model calculates power output based on various physiological parameters and running time.

```{r}
pmax <- 25    # Maximal instantaneous power (W)
wprime <- 20000  # Anaerobic energy reserve (J)
cp <- 300     # Critical power (W)
a <- 0.05     # Power index for durations > 30 minutes
time <- 1800  # Running time (s)

power <- OmPD(Pmax = pmax, Wprime = wprime, CP = cp, A = a, Time = time)
print(paste("OmniPD Power Output:", round(power, 2), "W/kg"))

# Compare power output for different durations
short_duration <- OmPD(Pmax = pmax, Wprime = wprime, CP = cp, A = a, Time = 1000)
long_duration <- OmPD(Pmax = pmax, Wprime = wprime, CP = cp, A = a, Time = 3600)
print(paste("Power output (short duration):", round(short_duration, 2), "W/kg"))
print(paste("Power output (long duration):", round(long_duration, 2), "W/kg"))

```

## Roy and Joyner Running Speed Model
The Roy and Joyner model estimates world record running speeds based on distance and various physiological parameters.

```{r}
s <- 0.1    # Aerobic metabolic slowdown index (m/s)
v0 <- 12    # Extrapolated speed at 100 m (m/s)
b <- 2      # Anaerobic boost index at 100 m (m/s)
beta <- 1.5 # Shape parameter for anaerobic boost curve
d <- 200    # Rate of decline of anaerobic boost (m)
distance <- 1000  # Running distance (m)

speed <- model_roy(S = s, V0 = v0, B = b, b = beta, d = d, Distance = distance)
print(paste("Roy and Joyner Model Speed:", round(speed, 2), "m/s"))

# Compare speeds for different distances
speed_100m <- model_roy(S = s, V0 = v0, B = b, b = beta, d = d, Distance = 100)
speed_marathon <- model_roy(S = s, V0 = v0, B = b, b = beta, d = d, Distance = 42195)
print(paste("Estimated speed (100m):", round(speed_100m, 2), "m/s"))
print(paste("Estimated speed (marathon):", round(speed_marathon, 2), "m/s"))

```



## Di Prampero Running Power
The di_prampero_running function converts running velocity to power output.

```{r}
# Example usage
power_output_dp <- di_prampero_running(
  velocity = 5,  # Running velocity in m/s
  distance = 1000  # Running distance in meters
)

print(paste("Estimated power output (Di Prampero model):", round(power_output_dp, 2), "W"))

```

## Working with World Records Data
ModPerformR includes datasets of world records for male and female athletes. Let's explore this data:

```{r}
# Load the datasets
data(WR_males)
data(WR_females)

# View the first few rows of male world records
head(WR_males)

# Calculate average world record times for males and females at 100m
avg_100m_male <- mean(WR_males$Time[WR_males$Distance == 100], na.rm = TRUE)
avg_100m_female <- mean(WR_females$Time[WR_females$Distance == 100], na.rm = TRUE)

print(paste("Average 100m world record time (males):", round(avg_100m_male, 2), "seconds"))
print(paste("Average 100m world record time (females):", round(avg_100m_female, 2), "seconds"))

```

## Conclusion

This vignette provides an introduction to the main functions and datasets in the ModPerformR package. These tools can be used to analyze running performance, estimate power output, and work with historical world record data.

