---
title: "running-power-calculations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{running-power-calculations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ModPerformR)
```

## Introduction
This vignette demonstrates how to use the `di_prampero_running` and `inv_di_prampero_running` functions from the ModPerformR package to calculate running power and estimate running distance based on power output.


## Calculating Running Power

The `di_prampero_running` function calculates the power output for a given running velocity and distance. Here's an example:

```{r}
velocity <- 5  # m/s
distance <- 5000  # meters

power <- di_prampero_running(velocity, distance)
print(paste("Estimated power output:", round(power, 2), "watts"))

```

We can also use custom parameters:

```{r}
power_custom <- di_prampero_running(
  velocity = 6,
  distance = 10000,
  BM = 65,  # Body mass in kg
  BSA = 1.7  # Body surface area in m^2
)
print(paste("Estimated power output (custom parameters):", round(power_custom, 2), "watts"))

```

## Estimating Running Distance
The `inv_di_prampero_running` function estimates the running distance for a given power output and duration. Here's an example:

```{r}
power <- 300  # watts
time <- 1800  # seconds (30 minutes)

distance <- inv_di_prampero_running(power, time)
print(paste("Estimated distance:", round(distance, 2), "meters"))

```

We can also use custom parameters:

```{r}
distance_custom <- inv_di_prampero_running(
  Power = 350,
  t = 3600,  # 1 hour
  BM = 75,  # Body mass in kg
  BSA = 1.9  # Body surface area in m^2
)
print(paste("Estimated distance (custom parameters):", round(distance_custom, 2), "meters"))

```

## Comparing Power Outputs at Different Velocities
Let's compare power outputs at different velocities for a 5000m run:

```{r}
velocities <- seq(3, 7, by = 0.5)
powers <- sapply(velocities, function(v) di_prampero_running(v, 5000))

# Create a data frame for easy viewing
comparison <- data.frame(
  Velocity = velocities,
  Power = powers
)

print(comparison)

```

## Estimating Distances for Different Power Outputs
Now, let's estimate the distances covered in a 30-minute run at different power outputs:

```{r}
powers <- seq(200, 400, by = 50)
distances <- sapply(powers, function(p) inv_di_prampero_running(p, 1800))

# Create a data frame for easy viewing
comparison <- data.frame(
  Power = powers,
  Distance = distances
)

print(comparison)

```



